---
title: "Collaborative analysis of gun legislation time series"
author: "BST 260 students"
name: "Jingyuan Xiao"
---

# Data acquisition

Grab the gun legislation data (originally from <https://www.rand.org/pubs/tools/TLA243-2-v4.html>).

```{r getd,message=FALSE}
if (!requireNamespace("readxl")) install.packages("readxl")
library(readxl)
if (!exists("lawdat")) {
  download.file("https://tinyurl.com/randgunlaw", (tf <- tempfile(fileext = ".xlsx")),mode = "wb")
  lawdat = read_xlsx(tf, 2)
}
```

# Data massaging for ease of use

Rename the variables for convenience.

```{r lk1,message=FALSE}
library(dplyr)
nd = lawdat |> rename( id=`Law ID`, state=State, 
  abbr=`State Postal Abbreviation`, fips=`FIPS Code`, 
  numclass=`Law Class (num)`, lawclass=`Law Class`, lawclasssub=`Law Class Subtype`, 
  handorlong=`Handguns or Long Guns`, effect=`Effect`, 
  changetype=`Type of Change`, edatenote=`Effective Date Note`, 
  effyear=`Effective Date Year`, effmonth=`Effective Date Month`, 
  effday=`Effective Date Day`, citation=`Statutory Citation`, 
  content=`Content`, supyear=`Supersession Date Year`, 
  supmonth=`Supersession Date Month`, supday=`Supersession Date Day`, 
  ageformin=`Age for Minimum Age Laws`, 
  handwaitdays=`Length of Waiting Period (days, handguns)`, 
  addit=`Additional Context and Notes`, 
  discrep=`Discrepancies`)
```

Use lubridate to create `effdate`, a Date-valued column.

```{r dodate,message=FALSE}
library(lubridate)
nd = nd |> mutate(effdate=ymd(paste(effyear, effmonth, effday, sep="-")))
```

# Data visualization

## Static

Visualize the big picture of counts of legislative events classified as Restrictive or Permissive after 1985.

```{r lkcounts,message=FALSE}
library(ggplot2)
ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01")), 
   aes(x=effdate)) + geom_histogram(binwidth=365) + ylim(0,350) +
   ggtitle("Counts of restrictive legislative changes after 1985")
ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01")), 
   aes(x=effdate)) + geom_histogram(binwidth=365) + ylim(0,350) +
   ggtitle("Counts of permissive legislative changes after 1985")
```

## Interactive

Here we use plotly so that we can hover over elements of the plot to get more information.

```{r doply,message=FALSE}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01")), 
   aes(x=effdate, text=paste(state,lawclass))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes after 1985")
ggplotly(g1)
```

# Per-state sections

## Alabama

```{r somecode,echo=FALSE}
# this code written by claude opus 5.5
wrap_hover <- function(x, width = 40, max_lines = Inf, sep = "<br>", escape = TRUE) {
  x <- as.character(x)
  vapply(x, function(s) {
    if (is.na(s) || !nzchar(s)) return(s)
    if (escape) {                       # keep "<", ">", "&" from being read as HTML
      s <- gsub("&", "&amp;", s, fixed = TRUE)
      s <- gsub("<", "&lt;",  s, fixed = TRUE)
      s <- gsub(">", "&gt;",  s, fixed = TRUE)
    }
    # split any word longer than the line width (long IDs, snake_case names, URLs)
    s <- gsub(sprintf("(\\S{%d})(?=\\S)", width - 1), "\\1 ", s, perl = TRUE)
    lines <- strwrap(s, width = width)
    if (length(lines) > max_lines) {
      lines <- lines[seq_len(max_lines)]
      lines[max_lines] <- paste0(lines[max_lines], " \u2026")
    }
    paste(lines, collapse = sep)
  }, character(1), USE.NAMES = FALSE)
}
```

```{r doplyAL}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Alabama"), 
   aes(x=effdate, text=paste(state,lawclass,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Alabama after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Alaska

```{r doplyAK}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Alaska"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Alaska after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
library(plotly)
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Alaska"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Alaska after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Arizona

Taehwan (Tae) Kwak

Github handle: tmk9872

```{r doplyAZ}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Arizona"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Arizona after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Arizona"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Arizona after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Arkansas - <https://github.com/AustinCo03>
```{r doplyAR}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Arkansas"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Arkansas after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
library(plotly)
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Arkansas"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Arkansas after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## California

```{r doplyCA}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="California"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in California after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="California"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in California after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Colorado

## Connecticut

GitHub: Qinyu57
Name: Qinyu Xu
```{r}
library(plotly)
#GitHub: Qinyu57
#Name: Qinyu Xu
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Connecticut"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Connecticut after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Connecticut"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Connecticut after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## Delaware

## Florida

## Georgia

## Hawaii

## Idaho

## Illinois
```{r illinois-restrictive, message=FALSE}
library(plotly)

g_il_restrictive = ggplot(
  nd |> filter(
    effect == "Restrictive",
    effdate > ymd("1985-01-01"),
    state == "Illinois"
  ),
  aes(
    x = effdate,
    text = paste(state, lawclass, changetype, wrap_hover(content))
  )
) +
  geom_histogram(binwidth = 365) +
  ggtitle("Counts of restrictive legislative changes in Illinois after 1985")

ggplotly(g_il_restrictive) |>
  plotly::style(hoverlabel = list(align = "left"))
```

```{r illinois-permissive, message=FALSE}
g_il_permissive = ggplot(
  nd |> filter(
    effect == "Permissive",
    effdate > ymd("1985-01-01"),
    state == "Illinois"
  ),
  aes(
    x = effdate,
    text = paste(state, lawclass, changetype, wrap_hover(content))
  )
) +
  geom_histogram(binwidth = 365) +
  ggtitle("Counts of permissive legislative changes in Illinois after 1985")

ggplotly(g_il_permissive) |>
  plotly::style(hoverlabel = list(align = "left"))
```
## Indiana

## Iowa

github/LeoMx11

```{r}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Iowa"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Iowa after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Iowa"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Iowa after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Kansas

## Kentucky

```{r doplyKY}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Kentucky"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Kentucky after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Kentucky"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Kentucky after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Louisiana
```{r}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Louisiana"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Louisiana after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
library(plotly)
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Louisiana"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Louisiana after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## Maine

```{r doplyME}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Maine"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Maine after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Maine"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Maine after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## Maryland

## Massachusetts

## Michigan

```{r doplyMI}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Michigan"), 
   aes(x=effdate, text=paste(state,lawclass,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Michigan after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))

g2 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Michigan"), 
   aes(x=effdate, text=paste(state,lawclass,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Michigan after 1985")
ggplotly(g2) |> plotly::style(hoverlabel = list(align = "left"))
```
## Minnesota
```{r doplyMN}
library(plotly)

g1 = ggplot(nd |> filter(
  effect=="Restrictive",
  effdate > ymd("1985-01-01"),
  state=="Minnesota"),
  aes(x=effdate,
      text=paste(state,lawclass,changetype,wrap_hover(content)))) +
  geom_histogram(binwidth=365) +
  ggtitle("Counts of restrictive legislative changes in Minnesota after 1985")

ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))

g1 = ggplot(nd |> filter(
  effect=="Permissive",
  effdate > ymd("1985-01-01"),
  state=="Minnesota"),
  aes(x=effdate,
      text=paste(state,lawclass,changetype,wrap_hover(content)))) +
  geom_histogram(binwidth=365) +
  ggtitle("Counts of permissive legislative changes in Minnesota after 1985")

ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## Mississippi

## Missouri

## Montana

## Nebraska

## Nevada

## New Hampshire

## New Jersey
```{r}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="New Jersey"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in New Jersey after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="New Jersey"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in New Jersey after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## New Mexico
```{r}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="New Mexico"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in New Mexico after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))

g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="New Mexico"),
  aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
  ggtitle("Counts of permissive legislative changes in New Mexico after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```


## New York
```{r doplyNY}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="New York"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in New York after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="New York"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in New York after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```

## North Carolina
```{r doplyNC}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="North Carolina"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in North Carolina after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))

g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="North Carolina"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in North Carolina after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## North Dakota
```{r doplyND}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="North Dakota"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in North Dakota after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="North Dakota"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in North Dakota after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## Ohio

## Oklahoma
```{r doplyOK}
library(plotly)
g1 = ggplot(nd |> filter(effect=="Restrictive", effdate > ymd("1985-01-01"), state=="Oklahoma"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of restrictive legislative changes in Oklahoma after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
g1 = ggplot(nd |> filter(effect=="Permissive", effdate > ymd("1985-01-01"), state=="Oklahoma"), 
   aes(x=effdate, text=paste(state,lawclass,changetype,wrap_hover(content)))) + geom_histogram(binwidth=365) +
   ggtitle("Counts of permissive legislative changes in Oklahoma after 1985")
ggplotly(g1) |> plotly::style(hoverlabel = list(align = "left"))
```
## Oregon
```{r}
data = nd %>% filter(effect %in% c("Restrictive", "Permissive"), effdate > ymd("1985-01-01"),state == "Oregon")%>%mutate(year = year(effdate))
graph = ggplot(data, aes(x = factor(year), fill = effect)) + geom_bar(position = "dodge") +
  labs(
    title = "Legislative changes in Oregon after 1985",
    x = "Year",
    y = "Count",
    fill = "Effect")
ggplotly(graph) |> plotly::style(hoverlabel = list(align = "left"))

```
## Pennsylvania

## Rhode Island

## South Carolina

## South Dakota

## Tennessee

## Texas

## Utah

## Vermont

## Virginia

## Washington

## West Virginia

## Wisconsin

## Wyoming
```{r doplyWY}
library(plotly)

g1 = ggplot(
  nd |> filter(
    effect == "Restrictive",
    effdate > ymd("1985-01-01"),
    state == "Wyoming"
  ),
  aes(
    x = effdate,
    text = paste(state, lawclass, changetype, wrap_hover(content))
  )
) +
  geom_histogram(binwidth = 365) +
  ggtitle("Counts of restrictive legislative changes in Wyoming after 1985")

ggplotly(g1) |> 
  plotly::style(hoverlabel = list(align = "left"))


g1 = ggplot(
  nd |> filter(
    effect == "Permissive",
    effdate > ymd("1985-01-01"),
    state == "Wyoming"
  ),
  aes(
    x = effdate,
    text = paste(state, lawclass, changetype, wrap_hover(content))
  )
) +
  geom_histogram(binwidth = 365) +
  ggtitle("Counts of permissive legislative changes in Wyoming after 1985")

ggplotly(g1) |> 
  plotly::style(hoverlabel = list(align = "left"))
```




